---
title: "Multi Linear Regression"
output: html_document
date: "2025-05-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r cars}
library(ggplot2)
```

## Including Plots

```{r }
# Load the CSV file
df <- read.csv("dv_socioeconomic_dataset_clipped.csv", stringsAsFactors = FALSE)

# Show number of rows (sample size)
nrow(df)

```
## 2. Fit a Multiple Linear Regression
```{r }
# Run linear regression model
fit <- lm(Domestic_violence_incide ~ Socioeconomic_index +
            Income + Education_years + Employment_status + Urban_status,
          data = df)

# Show model summary
summary(fit)

```
## 3. Interpretation (in Simple Terms)
From the summary above:

1. The intercept is the predicted DV when all values are zero.
2. SES Index: Positive effect; more SES → more DV, statistically significant.
3. Income: Negative effect; more income → slightly less DV, significant.
4. EducationYears: Dropped due to high overlap with other variables (collinearity).
5. EmploymentStatus: Negative effect, but not significant.
6. UrbanStatus: Slightly increases DV, somewhat significant.

```{r }
# Show just coefficients
coef(fit)
```
## 4. Model Deviation
We used all available variables. However, EducationYears was dropped due to collinearity (overlapping with other variables). This may lead to missing some important relationships and bias the model slightly. Also, the R-squared is low, so many other factors might be influencing DV.


## 5. Diagnostic: Check if Residuals are Normal (Shapiro-Wilk Test)
```{r }
# Get residuals
resid_vals <- resid(fit)

# Run the Shapiro-Wilk test
shapiro.test(resid_vals)

```
## 6. Plot: Residuals vs. Fitted Values
This plot helps us check if our model fits well. Ideally, the points should be randomly scattered.

```{r }
# Get fitted values
fitted_vals <- fitted(fit)

# Plot residuals
ggplot(data.frame(Fitted = fitted_vals, Residuals = resid_vals),
       aes(x = Fitted, y = Residuals)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Residuals vs Fitted",
       x     = "Fitted Values",
       y     = "Residuals") +
  theme_minimal()

```

